---
title: "Assignment B1"
author: "Jessica Chalissery"
date: "2023-10-26"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Welcome! 

This is my assignment B1 which is focused on learning how to write a function. The function I made is something that would have been useful to use during the mini data analysis projects we had to complete during STAT545A when I was looking into the *vancouver_trees* dataset from the *datateachr* package. This function is documented using *roxygen2* formatting.  


### Packages Needed

Here are the packages you will need to load in order to use this function and run the examples/tests. 
```{r}
library(datateachr)
library(dplyr)
library(roxygen2)
library(palmerpenguins)
library(testthat)
```

### Function

I chose to name this function *average_values_by_group*, since it takes the average value of each group in a dataset. This is a function that is quite versatile. It can take different individual datasets and separate it into groups based on a column of your choosing, and then find the mean value of a chosen measure (another column) for each group. The column chosen for the measure must have numeric values. Note that this function ignores NA values from the column chosen for the measure parameter so that it can calculate the mean. 
```{r}

#' @title Average Values By Group Function
#' 
#' @description This function takes a dataset and groups it by a chosen variable, then finds the mean value of a chosen measure for each group. 
#' 
#' @param dataset is the parameter for the **dataset** you are inputting into the function.
#' @param group is the parameter for the variable you choose to **group** the dataset by. This variable should be a column name from the dataset. 
#' @param measure is the parameter for the **measure** you choose to find the mean value of for each of the groups. The column you choose for the measure parameter must have numeric values. 
#' 
#' @return This function returns a dataframe with each *group* in the first column, and the *mean value of the measure for each group* in the second column. 


average_values_by_group <- function(dataset, group, measure) {
  var <- dataset %>% pull({{measure}}) %>% is.numeric()
  if(!var) {
    stop("The 'measure' parameter must have numeric values.")
  }
  dataset %>% group_by({{group}}) %>% summarize(average = mean({{measure}}, na.rm = TRUE))
}


```

### Examples

Here I will demonstrate how to use the function - it is quite versatile! 

```{r}

#' @examples These are examples showing that this function works with multiple dataset, variable and measure inputs.

average_values_by_group(vancouver_trees, genus_name, diameter)

average_values_by_group(vancouver_trees, species_name, latitude) 

average_values_by_group(penguins, island, bill_length_mm)

```
### Testing
Here I will be testing whether the function works in specific situations.

The first test will check whether this function works when the measure has no NA values. In this case, we are checking that we can compare two values from the dataframe and pass the test.
```{r}
test_that("Test if this function works when the measure has no NA values", {
  sample_data1 <- data.frame(Group = c("A", "A", "B", "B", "B"), Value = c(10, 20, 30, 40, 50))
  result <- average_values_by_group(sample_data1, Group, Value)
  expect_gt(result$average[2], result$average[1])
})
```

The second test checks whether this function works when the measure has NA values. This function should ignore NA values in order to calculate the mean. Similar to the previous test, we are checking that we can compare two values from the dataframe and pass the test, even with NA values present.
```{r}
test_that("Test if this function works when the measure has NA values", {
  sample_data2 <- data.frame(Group = c("A", "A", "B", "B", "B"), Value = c(10, 20, NA, 40, 50))
  result <- average_values_by_group(sample_data2, Group, Value)
  expect_lt(result$average[1], result$average[2])
})
```
The third test checks whether this function works when the measure has string values instead of numeric values. Since our function should only work when the measure parameter has numeric values, we are checking to make sure that when there are non-numeric values in the measure parameter, it gives the appropriate error message.
```{r}
test_that("Test if this function works when the measure has no values", {
  sample_data3 <- data.frame(Group = c("A", "A", "B", "B", "B"), Value = c("one", "two", "three", "four", "five"))
  expect_error(average_values_by_group(sample_data3, Group, Value), "The 'measure' parameter must have numeric values.")
})
```
Hooray! The tests passed! We succeeded :)